Manuscript_draft

### 2020-12-17 manuscript_draft.Rmd
manuscript work

# Height model fit plot
htModelFit.plot<-
  X %>% 
  filter(is.na(DBT)) %>% # filter for trees with no broken tops
  mutate(h2=exp(fixef(htDiamFit)[1]+fixef(htDiamFit)[2]*log(DBH))) %>% 
  ggplot(aes(x=Height,y=h2))+
  geom_point()+
  xlim(0,40)+
  ylim(0,40)+
  geom_smooth(method="lm",se=F)+
  ylab("Predicted height (m) from diameter-height model")+
  xlab("Measured height (m)")

@@

----------------------------
```{r height model fit}

htModelFit.plot

@@

# diamClass.plot<-cowplot::plot_grid(diamVol.plot,diamBA.plot,diamK.plot,ncol=1)

@@

plotData.X<-
# Determine which plots have measurements in 1992 and 2019
  X %>% 
  #  filter(Year %in% c(1992,2019)) %>% 
  #  filter(Status=="Live") %>% 
  # 
  # group_by(Plot,Year) %>% 
  # summarise(n=n()) %>% 
  # ungroup() %>% 
  # group_by(Plot) %>% 
  # summarise(n=length(Year)) %>% 
  # filter(n>1) %>% # pick plots with measurements in both years
  # pull(Plot) %>% 
  # as.data.frame() %>% 
  # setNames("Plot") %>% 
  # 
  # # Now filter original dataset 
  # left_join(X,by="Plot") %>% 
  # mutate(Plot=fct_drop(Plot)) %>%  # drop unused levels
  # 
  # # Filter for years of interest
  # filter(Year %in% c(1992,2019)) %>% 
  # mutate(Year=fct_drop(Year)) %>% 
  filter(Status=="Live") %>%

@@

# PLotting

standK.plot<-
  standPlot("SPH",y.label="Tree dens. (sph)")+
  theme(axis.text.x=element_blank(),axis.title.x=element_blank())

# could put in standQMD.plot as well

@@

  # formatting
  rownames_to_column("Fixed effect") %>% 
  dplyr::select(Variable,`Fixed effect`,Sum.Sq:F.value,P.value="Pr..F.") %>% 
  mutate_at(vars("F.value","P.value"),round,5) %>% 
  mutate(`Fixed effect`=rep(c("Treatment","Year","Trt.*Year"),times=3)) %>% 
  rename(`Sum sq.`="Sum.Sq",
         `Mean sq.`="Mean.Sq",
         `Num. DF`="NumDF",
         `Den. DF`="DenDF",
         `F value`="F.value",
         `p value`="P.value")

@@

```{r fig.height=10}

standPlots
```

`r figNums("standPlots",caption="Mean plot attributes by treatment unit and year.  Whiskers are one standard error of the mean.  Vol., volume; BA, basal area; Tree dens., tree density; QMD, quadratic mean diameter")`

*Note that sample size varies with year.  Some plots were not measured in 1992.  This likely explains the dip in basal area and volume in TU:20 from 1992 to 1994, despite no recorded mortality in 1992/1994.*

@@

```{r summarize changes in species composition}

# create stand-level plots by diameter distribution
sppVol.plot<-sppComp.plot("BA")

figNums("sppVol.plot",display=FALSE)

# no need to provide other figures, as they're very similar

@@

A comparison of proportional species composition by diameter class, treatment and year showed that fir and spruce were organized similarly between the high RBA and control units (`r figNums("sppVol.plot",display="cite")`).  In 1992, the largest diameter classes were dominated by spruce, whereas fir was the dominant species in the smallest diameter classes.  This pattern was found for volume, basal area, and tree density.  We only show volume in this document.  By 2019, the fir-spruce composition by distribution class was maintained in these two units, with small increases in spruce composition in the smallest diameter class.

@@

```{r species composition plot}
sppVol.plot

`r figNums("sppVol.plot",caption="Spruce-fir volume proportional composition by diameter class, treatment and year.")`

*Note this figure is almost identical if we compare stems, basal area or volume. For brevity, we only show volume here.*

@@

`r tablesNums("fitBA.table",caption="Summary of mixed effects model of tree-level basal area increment")`

@@

`r tablesNums("fitBApct.table",caption="Summary of mixed effects model of tree-level basal area increment as percentage of initial tree size")`

@@

@@

>>>>

@@

@@

# diamClass.plot<-cowplot::plot_grid(diamVol.plot,diamBA.plot,diamK.plot,ncol=1)
```

@@

plotData.X<-
# Determine which plots have measurements in 1992 and 2019
  X %>%

  filter(Status=="Live") %>%

@@

@@

  # formatting
  rownames_to_column("Fixed effect") %>% 
  dplyr::select(Variable,`Fixed effect`,Sum.Sq:F.value,P.value="Pr..F.") %>% 
  mutate_at(vars("F.value","P.value"),round,5) %>% 
  mutate(`Fixed effect`=rep(c("Treatment","Year","Trt*Year"),times=3)) %>% 
  rename(`Sum sq.`="Sum.Sq",
         `Mean sq.`="Mean.Sq",
         `Num. DF`="NumDF",
         `Den. DF`="DenDF",
         `F value`="F.value",
         `p value`="P.value")

@@

@@

```{r caption for species composition plots}

figNums("sppComp.plot",display=FALSE)

@@

A comparison of proportional species composition by diameter class, treatment and year showed that fir and spruce were organized similarly between the high RBA and control units (`r figNums("sppComp.plot",display="cite")`).  In 1992, the largest diameter classes were dominated by spruce, whereas fir was the dominant species in the smallest diameter classes.  This pattern was found for volume, basal area, and tree density.  We only show volume in this document.  By 2019, the fir-spruce composition by distribution class was maintained in these two units, with small increases in spruce composition in the smallest diameter class.

@@

@@

`r tableNums("fitBA.table",caption="Summary of mixed effects model of tree-level basal area increment")`

@@

`r tableNums("fitBApct.table",caption="Summary of mixed effects model of tree-level basal area increment as percentage of initial tree size")`

@@

```{r function to create figure of plot means, fig.height=10}

# Plotting

standK.plot<-
  standPlot("SPH",y.label="Tree dens. (sph)")

# display plot
standPlots
```

`r figNums("standPlots",caption="Mean plot attributes by treatment unit and year.  Whiskers are one standard error of the mean.  Vol., volume; BA, basal area; Tree dens., tree density")`

```{r species composition plots by diameter class}

# create stand-level plots by diameter distribution
sppComp.plot("BA")

`r figNums("sppComp.plot",caption="Spruce-fir volume proportional composition by diameter class, treatment and year.")`

*Note this figure is almost identical for compairsons of basal area, volume or tree density. For brevity, we only show basal area.*

@@

```{r height model fit}

# Height model fit plot

  X %>% 
  filter(is.na(DBT)) %>% # filter for trees with no broken tops
  mutate(h2=exp(fixef(htDiamFit)[1]+fixef(htDiamFit)[2]*log(DBH))) %>% 
  ggplot(aes(x=Height,y=h2))+
  geom_point()+
  xlim(0,40)+
  ylim(0,40)+
  geom_smooth(method="lm",se=F)+
  ylab("Predicted height (m) from diameter-height model")+
  xlab("Measured height (m)")

# References
